---
title: "2.02_PCA-Example"
author:
- Andrew Davis, PhD, Unilever Human Biology
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::html_document:
    toc: TRUE
---

```{r, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 comment = "#>",
 out.width = "100%"
)
```

## Objectives:

This script presents a simple script for plotting a PCA of the mtcars dataset, which comes preloaded into R. This dataset consists of data on 32 models of car, taken from an American motoring magazine (1974 Motor Trend magazine).

## Install/Load dependencies

```{r, echo=TRUE, message=FALSE, Loading_Dependencies_1}
# Load dependencies:
library(ggfortify)
```

## Perform Principal Component Analysis:

```{r compute_pca}
mtcars.pca <- prcomp(mtcars, center = TRUE,scale. = TRUE)
summary(mtcars.pca)
```

## Plot PCA:

```{r}
autoplot(mtcars.pca, data = mtcars, colour = 'cyl')
```

## Plot PCA with labeled eigenvectors

```{r}
pdf("../Results/PCA_Eigenvectors.pdf")
autoplot(mtcars.pca, data = mtcars, colour = 'cyl',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3)
dev.off()

autoplot(mtcars.pca, data = mtcars, colour = 'cyl',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3)
```

### Session information

```{r session_info_202}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.
